---
title: "Chapter_5"
author: "Alper"
date: "2022-11-18"
output: html_document
---

# Table 1 demographics
## adults vs children
```{r}
library(gtsummary)
tabledata1<- flowsom_df_wide %>% 
  filter(comp1_adultsvschildren_pos %in% "TRUE") %>%
  filter(adult_child %!in% "adult_control") %>%
  select(
    comp1_adultsvschildren_pos,comp2_mildvsmoderate_adult,comp3_PRR,
    timepoint, PCR_test, PCR_progress,severity_text, adult_child, sex, age,
    contains(c("CD4_","CD8_","B_","Mono_","NK_"))) %>%
  # na.omit() %>%
  select(
    timepoint,PCR_test,PCR_progress,severity_text, adult_child, sex, age) %>%
  mutate_at(c("severity_text"), list(~as.factor(.))) %>% 
  droplevels()

tabledata1 %>% 
  filter(adult_child %in% "adult") %>%
  tbl_summary(by = c(severity_text))


tabledata1 %>% 
  filter(adult_child %in% "child") %>%
  tbl_summary(by = c(severity_text))

tabledata1 %>%   tbl_summary(by = c(severity_text))



# disease onset 
flowsom_df_wide %>% 
  full_join(ffx_onset, by="subject_identifier") %>%
  filter(comp1_adultsvschildren_pos %in% "TRUE") %>%
  filter(adult_child %!in% "adult_control") %>%
  select(subject_identifier,T1_onset,
    comp1_adultsvschildren_pos,comp2_mildvsmoderate_adult,comp3_PRR,
    timepoint, PCR_test, PCR_progress,severity_text, adult_child, sex, age,
    contains(c("CD4_","CD8_","B_","Mono_","NK_"))) %>%
  group_by(adult_child) %>%
  summarise(
    median_T1_onset = median(T1_onset, na.rm = TRUE),
    Q1_T1_onset = quantile(T1_onset, 0.25, na.rm = TRUE),
    Q3_T1_onset = quantile(T1_onset, 0.75, na.rm = TRUE),
    IQR_T1_onset = IQR(T1_onset, na.rm = TRUE)
  )
```
### age distribution in children and adults
```{r}
tabledata1 %>% 
  # filter(adult_child %in% "child") %>%
  ggplot(aes(x=adult_child, y=age))+
  geom_boxplot(aes(fill=adult_child), alpha=0.9,outlier.size=0,outlier.colour="white")+
  geom_jitter(aes(fill=adult_child), alpha=0.5, width = 0.3, shape=21,size=1.5) +
  theme_classic()+
  theme(legend.position = "none")+
  xlab("")+
  ylab("Age (years)")


ggsave(file = paste0(AvsC_save.dir, "adult_vs_children_age.png", sep=""), width =2 , height =3)

tabledata1 %>% 
  filter(adult_child %in% "child") %>%
  ggplot(aes(x=adult_child, y=age))+
  geom_boxplot(aes(fill=adult_child), alpha=0.9,outlier.size=0,outlier.colour="white")+
  geom_jitter(aes(fill=adult_child), alpha=0.5, width = 0.3, shape=21,size=1.5) +
  theme_classic()+
  theme(legend.position = "none")+
  xlab("")+
  ylab("Age (years)")+
  scale_y_continuous(breaks=seq(0,18,1))


ggsave(file = paste0(AvsC_save.dir, "children_age.png", sep=""), width =2 , height =3)
```

## disease severity
```{r}
tabledata2<- flowsom_df_wide %>% 
  filter(comp2_mildvsmoderate_adult %in% "TRUE") %>%
  select(
    comp1_adultsvschildren_pos,comp2_mildvsmoderate_adult,comp3_PRR,
    timepoint, PCR_test, PCR_progress,severity_text, adult_child, sex, age,
    contains(c("CD4_","CD8_","B_","Mono_","NK_")))%>%  
  filter(severity_text %!in% "none")%>%
  # na.omit() %>%
  select(
    timepoint,PCR_test,PCR_progress,severity_text, adult_child, sex, age) %>%
  mutate_at(c("severity_text"), list(~as.factor(.))) %>% 
  droplevels()

tabledata2 %>% 
  filter(adult_child %in% "adult") %>%
  tbl_summary(by = c(severity_text))

df <- flowsom_df_wide %>% 
  filter(comp2_mildvsmoderate_adult %in% "TRUE") %>%
  select(subject_identifier,
    comp1_adultsvschildren_pos,comp2_mildvsmoderate_adult,comp3_PRR,
    timepoint, PCR_test, PCR_progress,severity_text, adult_child, sex, age,
    contains(c("CD4_","CD8_","B_","Mono_","NK_")))%>%  
  filter(severity_text %!in% "none")%>%
  # na.omit() %>%
  select(subject_identifier,
    timepoint,PCR_test,PCR_progress,severity_text, adult_child, sex, age) %>%
  mutate_at(c("severity_text"), list(~as.factor(.))) %>% 
  droplevels() 

df[duplicated(df$subject_identifier) | duplicated(df$subject_identifier, fromLast = TRUE), ]

```

## Recovering cases
```{r}
tabledata3<- flowsom_df_wide %>% 
  filter(comp3_PRR %in% "TRUE") %>%
  select(
    comp1_adultsvschildren_pos,comp2_mildvsmoderate_adult,comp3_PRR,
    timepoint, PCR_test, PCR_progress,severity_text, adult_child, sex, age,
    contains(c("CD4_","CD8_","B_","Mono_","NK_")))%>%  
  filter(severity_text %!in% "none")%>%
  select(
    timepoint,PCR_test,PCR_progress,severity_text, adult_child, sex, age) %>%
  mutate(severity_text = if_else(severity_text == "none", "mild", severity_text)) %>%
  mutate_at(c("severity_text"), list(~as.factor(.))) %>%
  filter(timepoint %in% 1)

tabledata3 %>% 
  filter(adult_child %in% "adult") %>%
  tbl_summary(by = c(severity_text))
```

# Onset of symptoms and IQR
```{r}
ffx_onset %>%
  filter(subject_identifier %in% subject_IDs_FFX) %>%
  left_join(ffx_subject[c(2,8,11,12,13)]) %>%
  group_by(adult_child) %>%
  summarise(
    median_T1_onset = median(T1_onset, na.rm = TRUE),
    Q1_T1_onset = quantile(T1_onset, 0.25, na.rm = TRUE),
    Q3_T1_onset = quantile(T1_onset, 0.75, na.rm = TRUE),
    IQR_T1_onset = IQR(T1_onset, na.rm = TRUE)
  )

```




# PCA for PCR+ adults vs children
## data organization
```{r}
pca_immune_comp1 <- flowsom_df_wide  %>%
  filter(comp1_adultsvschildren_pos %in% "TRUE") %>%
  filter(adult_child %!in% "adult_control") %>%
  dplyr::select(sample_identifier,contains(c("CD4_","CD8_","B_","Mono_","NK_"))) %>%
  column_to_rownames(var = "sample_identifier") %>%
  as.data.frame() %>%
  na.omit()


pca_subject_comp1 <- flowsom_df_wide  %>%
  filter(comp1_adultsvschildren_pos %in% "TRUE") %>%
  dplyr::select(sample_identifier, severity_text, severity_merged, sex, adult_child) %>%
  filter(adult_child %!in% "adult_control") %>%
  filter(sample_identifier %in% rownames(pca_immune_comp1)) %>%
  column_to_rownames(var = "sample_identifier") %>%
  as.data.frame()

# timepoints of between first PCR+ sample
## 1
flowsom_df_wide  %>%
  filter(comp1_adultsvschildren_pos %in% "TRUE") %>% 
  filter(timepoint %in% 1)
## 2
flowsom_df_wide  %>%
  filter(comp1_adultsvschildren_pos %in% "TRUE") %>% 
  filter(timepoint %in% 2)
```

## plot PCAs
```{r}
## run PCA
res.pca <- PCA(pca_immune_comp1, graph = TRUE) #dat is scaled

## extract eigen values
eig.val <- get_eigenvalue(res.pca) 

## Screeplot visualise eigen values
PCA_screeplot <- fviz_eig(res.pca, addlabels = TRUE, ylim = c(0, 25)) 
pdf(file = paste0(PCA_save.dir, "PCA_screeplot_AvsC.pdf", sep=""), width =6 , height =5) ; PCA_screeplot ; dev.off()

## Extract the results for individuals only
var <- get_pca_var(res.pca)
head(var$coord) #coordinates
head(var$cos2) #Cos2: quality on the factore map
head(var$contrib) #Contributions to the principal components

x1 <-var$contrib
x2 <-var$cos2

## Extract the results for variables only
ind <- get_pca_ind(res.pca)
# head(ind$coord) #coordinates
# head(ind$cos2) #Cos2: quality on the factore map
# head(ind$contrib) #Contributions to the principal components


## Generate PCA d7d0 predictors of HI
## PCA, generating plots 
## Visualization of PCA plot with geom density 
pca_geomdens <- fviz_pca_ind(res.pca,
                             geom.ind = "point", # show points only (but not "text")
                             legend.title = "PCA density") + 
geom_density_2d()

## save
pdf(file = paste0(PCA_save.dir, "PCA_density_AvsC.pdf", sep=""), width =5 , height =5) ; pca_geomdens ; dev.off()

## Visualization of top 10 parameters,  cos2
PCA_variable_cos2_topx <- fviz_pca_var(res.pca, 
                                       col.var = "cos2",
                                       gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
                                       repel=TRUE,
                                       title ="",
                                       select.var = list(name = NULL, cos2 = 10, contrib = NULL))+theme_classic()
## save
pdf(file = paste0(PCA_save.dir, "PCA_variable_cos2_topx_AvsC.pdf", sep=""), width =5 , height =4) ; PCA_variable_cos2_topx ; dev.off()


## Visualization of cos2 all
PCA_variable_cos2 <- fviz_pca_var(res.pca, 
                                  col.var = "cos2",
                                  gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
                                  repel=TRUE,
                                  alpha.var = "cos2",
                                  title ="cos2 to PCs")+theme_classic()
## save
pdf(file = paste0(PCA_save.dir, "PCA_variable_cos2_AvsC.pdf", sep=""), width =10 , height =10) ; PCA_variable_cos2 ; dev.off()


## PCA with PCR+ adults, children 
pdf(file = paste0(PCA_save.dir,"PCA_AvsC_adults_children.pdf", sep=""), width =3.5 , height =3)
fviz_pca_ind(res.pca,
             geom.ind = "point", # show points only (but not "text")
             habillage = as.factor(pca_subject_comp1$adult_child),
             palette = c("#77AADD","#BBCC33"),
             ellipse.type = c("t"),
             addEllipses = TRUE,
             pointsize = 0.9,
             alpha.ind = 1
             )+
  theme_classic()
dev.off()

## location of individuals
pdf(file = paste0(PCA_save.dir,"PCA_individuals_AvsC.pdf", sep=""), width =7 , height =7)
fviz_pca_ind(res.pca,
             geom.ind = "text", # show points only (but not "text")
             gradient.cols = heat.colors(60),
             legend.title = "")
dev.off()

```

# PCA for PCR+ adults, mild vs moderate
## data organization
```{r}
pca_immune_comp2 <- flowsom_df_wide  %>%
  filter(comp2_mildvsmoderate_adult %in% "TRUE") %>%
  dplyr::select(sample_identifier,contains(c("CD4_","CD8_","B_","Mono_","NK_"))) %>%
  column_to_rownames(var = "sample_identifier") %>%
  data.frame() %>%
  na.omit()

pca_subject_comp2 <- flowsom_df_wide  %>%
  filter(comp2_mildvsmoderate_adult %in% "TRUE") %>%
  dplyr::select(sample_identifier, severity_text, severity_merged, sex, adult_child) %>%
  filter(sample_identifier %in% rownames(pca_immune_comp2))%>%
  column_to_rownames(var = "sample_identifier") %>%
  as.data.frame()

# timepoints PCR+ sample
## 1
flowsom_df_wide  %>%
  filter(comp2_mildvsmoderate_adult %in% "TRUE") %>% 
  filter(timepoint %in% 1)

## 2
flowsom_df_wide  %>%
  filter(comp2_mildvsmoderate_adult %in% "TRUE") %>% 
  filter(timepoint %in% 2)


```

## plot PCA
```{r}
## run PCA
res.pca <- PCA(pca_immune_comp2, graph = TRUE) #dat is scaled

## extract eigen values
eig.val <- get_eigenvalue(res.pca) 

## Screeplot visualise eigen values
PCA_screeplot <- fviz_eig(res.pca, addlabels = TRUE, ylim = c(0, 25)) 
pdf(file = paste0(PCA_save.dir, "PCA_screeplot_MivsMo.pdf", sep=""), width =6 , height =5) ; PCA_screeplot ; dev.off()

## Extract the results for individuals only
var <- get_pca_var(res.pca)
head(var$coord) #coordinates
head(var$cos2) #Cos2: quality on the factore map
head(var$contrib) #Contributions to the principal components

x1 <-var$contrib
x2 <-var$cos2

## Extract the results for variables only
ind <- get_pca_ind(res.pca)
# head(ind$coord) #coordinates
# head(ind$cos2) #Cos2: quality on the factore map
# head(ind$contrib) #Contributions to the principal components


## Generate PCA d7d0 predictors of HI
## PCA, generating plots 
## Visualization of PCA plot with geom density 
pca_geomdens <- fviz_pca_ind(res.pca,
                             geom.ind = "point", # show points only (but not "text")
                             legend.title = "PCA density") + 
geom_density_2d()

## save
pdf(file = paste0(PCA_save.dir, "PCA_density_MivsMo.pdf", sep=""), width =5 , height =5) ; pca_geomdens ; dev.off()

## Visualization of top 10 parameters,  cos2
PCA_variable_cos2_topx <- fviz_pca_var(res.pca, 
                                       col.var = "cos2",
                                       gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
                                       repel=TRUE,
                                       title ="",
                                       select.var = list(name = NULL, cos2 = 10, contrib = NULL))+theme_classic()
## save
pdf(file = paste0(PCA_save.dir, "PCA_variable_cos2_topx_MivsMo.pdf", sep=""), width =5 , height =4) ; PCA_variable_cos2_topx ; dev.off()


## Visualization of cos2 all
PCA_variable_cos2 <- fviz_pca_var(res.pca, 
                                  col.var = "cos2",
                                  gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
                                  repel=TRUE,
                                  alpha.var = "cos2",
                                  title ="cos2 to PCs")+theme_classic()
## save
pdf(file = paste0(PCA_save.dir, "PCA_variable_cos2_MivsMo.pdf", sep=""), width =10 , height =10) ; PCA_variable_cos2 ; dev.off()


## PCA with severity
pdf(file = paste0(PCA_save.dir,"PCA_MivsMo_severity_text.pdf", sep=""), width =4 , height =3)
fviz_pca_ind(res.pca,
             geom.ind = "point", # show points only (but not "text")
             habillage = as.factor(pca_subject_comp2$severity_text),
             palette = c("#44BB99","#77AADD", "#EE8866"),
             ellipse.type = c("t"),
             addEllipses = TRUE,
             pointsize = 0.9,
             alpha.ind = 1
             )+
  theme_classic()
dev.off()

## location of individuals
pdf(file = paste0(PCA_save.dir,"PCA_individuals_MivsMo.pdf", sep=""), width =7 , height =7)
fviz_pca_ind(res.pca,
             geom.ind = "text", # show points only (but not "text")
             gradient.cols = heat.colors(60),
             legend.title = "")
dev.off()

```
# Supplementary Table 1, Statistically different subsets between PCR+ adults and children
```{r}
df_stat <- flowsom_df_wide %>% 
  filter(comp1_adultsvschildren_pos %in% "TRUE") %>%
  filter(adult_child %!in% "adult_control") %>%
  dplyr::select(sample_identifier,adult_child,contains(c("CD4_","CD8_","B_","Mono_","NK_")))
  
## subject identifier and cluster number as.character
df_stat %<>% mutate_at(
  c("adult_child"),
  list(~as.character(.)))

## melt
df_stat_melt <- df_stat %>% 
  reshape2::melt(.)
df_stat_melt

### kruskal.test and Dunn's test
## With kruskal.test we identify overall which subsets are significant and then Dunn's test tells us which are different between pairs of clusters
kw.test.all <- df_stat_melt %>% 
  group_by(variable) %>% 
  kruskal_test(value ~ adult_child) %>% 
  adjust_pvalue(method = "BH") #BH correction

kw.test.all.eff.size <- df_stat_melt %>% #effect size
  group_by(variable) %>% 
  kruskal_effsize(value ~ adult_child) 

combine.kw <- kw.test.all %>% 
  left_join(kw.test.all.eff.size, by="variable")

## Number of variables with p.adj-value <0.05
length(combine.kw$variable)

## table summarizing effect size of the significant variables 
table(combine.kw$magnitude)

### Post-hoc analysis dunn's test, pairwise multiple comparisons to check which are significant in pairwise analysis.
dun.test.pos <- df_stat_melt %>% 
  group_by(variable) %>% 
  dunn_test(value ~ adult_child, p.adjust.method = "BH", detailed = TRUE)

head(dun.test.pos)
dun.test.pos_sub <- dun.test.pos %>% filter(p.adj<=0.05) #filter non significant data

dunntest_AvsC <- dun.test.pos_sub %>% arrange(-desc(p.adj))
writexl::write_xlsx(dunntest_AvsC, paste0(analysis.path, "/results/tables/dunn_AvsC.xlsx", sep=""))
```

# Statistically different subsets between PCR+ mild, moderate and severe adults
```{r}
df_stat <- flowsom_df_wide %>% 
  filter(comp2_mildvsmoderate_adult %in% "TRUE") %>%
  dplyr::select(sample_identifier,severity_text,contains(c("CD4_","CD8_","B_","Mono_","NK_"))) %>%
  dplyr::select(-Mono_12) %>%
  droplevels()

## subject identifier and cluster number as.character
df_stat %<>% mutate_at(
  c("severity_text"),
  list(~as.character(.)))

## melt
df_stat_melt <- df_stat %>% 
  reshape2::melt(.)
df_stat_melt

### kruskal.test and Dunn's test
## With kruskal.test we identify overall which subsets are significant and then Dunn's test tells us which are different between pairs of clusters
kw.test.all <- df_stat_melt %>% 
  group_by(variable) %>% 
  kruskal_test(value ~ severity_text) %>% 
  adjust_pvalue(method = "BH") #BH correction

kw.test.all.eff.size <- df_stat_melt %>% #effect size
  group_by(variable) %>% 
  kruskal_effsize(value ~ severity_text) 

combine.kw <- kw.test.all %>% 
  left_join(kw.test.all.eff.size, by="variable")

## Number of variables with p.adj-value <0.05
length(combine.kw$variable)

## table summarizing effect size of the significant variables 
table(combine.kw$magnitude)

### Post-hoc analysis dunn's test, pairwise multiple comparisons to check which are significant in pairwise analysis.
dun.test.pos <- df_stat_melt %>% 
  group_by(variable) %>% 
  do({
    print(paste("Processing:", .[["variable"]][1]))
    dunn_test(value ~ severity_text, data = ., p.adjust.method = "BH", detailed = TRUE)
  })


dun.test.pos_sub <- dun.test.pos %>% filter(p.adj<=0.05) #filter non significant data

dunntest_A_MvsM2 <- dun.test.pos_sub %>% arrange(-desc(p.adj))

writexl::write_xlsx(dunntest_A_MvsM2, paste0(analysis.path, "/results/tables/dunntest_A_MvsM2.xlsx", sep=""))
```


# Statistical analysis for PRR
```{r}
long_data <- flowsom_df_wide  %>%
  filter(comp3_PRR %in% "TRUE") %>%
  dplyr::select(subject_identifier,timepoint,contains(c("CD4_","CD8_","B_","Mono_","NK_"))) %>%
  as.data.frame() %>% 
  reshape2::melt(.)


# Loop through the immune variables and perform the analysis
immune_variables <- unique(long_data$variable)

# Initialize an empty data frame to store the results
anova_results <- data.frame(variable = character(),
                            timepoint_p_value = numeric(),
                            stringsAsFactors = FALSE)

# Loop through the immune variables and perform the analysis
for (var in immune_variables) {
  sub_data <- long_data %>% filter(variable == var)
  anova_result <- aov(value ~ timepoint + Error(subject_identifier/timepoint), data = sub_data)
  anova_summary <- summary(anova_result)
  
  # Extract the p-value for the timepoint factor
  timepoint_p_value <- anova_summary[[2]][[1]]["timepoint", "Pr(>F)"]

  # Append the results to the data frame
  anova_results <- anova_results %>%
    add_row(variable = var,
            timepoint_p_value = timepoint_p_value)
}

# Choose a significance level, e.g., 0.05
significance_level <- 0.05

# Filter for significant immune variables
significant_variables_PRR <- anova_results %>%
  filter(timepoint_p_value < significance_level)

```

# Plots PCR+ adults and children for significant subsets
## data prep
```{r echo=TRUE}
sub_df <- flowsom_df_wide  %>%
  filter(comp1_adultsvschildren_pos %in% "TRUE") %>%
  filter(adult_child %!in% "adult_control") %>%
  dplyr::select(sample_identifier,age,adult_child,contains(c("CD4_","CD8_","B_","Mono_","NK_"))) %>%
  droplevels() %>%
  as.data.frame()
```

## selection 
```{r echo=TRUE}

# Load the necessary libraries
library(ggpubr)

# Specify the names of the columns you want to loop through
vars_to_plot <- c("CD4_11","NK_17", "CD8_5", "NK_19", "B_7", "CD4_2", "B_3","NK_21")

# Initialize an empty list to store the plots
plot_list <- list()

# Loop through each variable and create a plot
for (var in vars_to_plot) {
  
  # Create the plot
  p <- ggplot(sub_df) +
    geom_point(
      aes(x = adult_child, y = !!sym(var), group = adult_child, color = adult_child),
      position = position_jitter(width = 0.1, seed = 42),
      size = 1,alpha=0.9,
      show.legend = FALSE
    ) +
    geom_half_violin(
      aes(x = adult_child, y = !!sym(var), group = adult_child, fill = adult_child),
      side = "r",
      position = position_nudge(0.13),
    ) +
    geom_half_boxplot(
      aes(x = adult_child, y = !!sym(var), group = adult_child, fill = adult_child),
      side = "l",
      outlier.shape = NA,
      center = FALSE,
      errorbar.draw = FALSE,
      width = .4,
      position = position_nudge(-0.13),
    ) +
    theme_classic() +
    stat_summary(aes(y = !!sym(var), x = adult_child), size = 0.2,
                 position = position_nudge(-0.23)) +
    scale_fill_manual(values = cols_c_AvsC) +
    scale_color_manual(values = cols_c_AvsC) +
    theme(legend.position = "none")+
    xlab("")
  
  # Append the plot to the list
  plot_list[[var]] <- p
}

# Arrange the plots in a grid and save as a single file
combined_plot <- do.call(ggarrange, c(plot_list, list(ncol = 4)))

ggsave(paste0(AvsC_save.dir, paste("AvsC_combined_plots.pdf"), sep=""),combined_plot, width = 8, height = 4)
       
```

# Plots mild vs moderate adults for significant subsets
## data prep 
```{r echo=TRUE}
sub_df <- flowsom_df_wide  %>%
  filter(comp2_mildvsmoderate_adult %in% "TRUE") %>%
  droplevels() %>%
  as.data.frame()
```

## selection
```{r echo=TRUE}
# Specify the names of the columns you want to loop through
vars_to_plot <- c("B_8" , "CD4_3", "Mono_4", "B_1", "B_9","B_19","NK_10","CD4_17")

# Initialize an empty list to store the plots
plot_list <- list()
plot_list_log <- list()

# Loop through each variable and create a plot
for (var in vars_to_plot) {
  
  # Create the plot
  p <- ggplot(sub_df) +
    geom_point(
      aes(x = severity_text, y = !!sym(var), group = severity_text, color = severity_text),
      position = position_jitter(width = 0.1, seed = 42),
      size = 1.5,
      show.legend = FALSE
    ) +
    geom_half_violin(
      aes(x = severity_text, y = !!sym(var), group = severity_text, fill = severity_text),
      side = "r",
      position = position_nudge(0.13),
    ) +
    geom_half_boxplot(
      aes(x = severity_text, y = !!sym(var), group = severity_text, fill = severity_text),
      side = "l",
      outlier.shape = NA,
      center = FALSE,
      errorbar.draw = FALSE,
      width = .4,
      position = position_nudge(-0.13),
    ) +
    theme_classic() +
    stat_summary(aes(y = !!sym(var), x = severity_text), size = 0.2,
                 position = position_nudge(-0.23)) +
    scale_fill_manual(values = cols_MivsMo2) +
    scale_color_manual(values = cols_MivsMo2) +
    theme(legend.position = "none")+
    xlab("")
  
  # Append the plot to the list
  plot_list[[var]] <- p
}

# Arrange the plots in a grid and save as a single file
combined_plot <- do.call(ggarrange, c(plot_list, list(ncol = 3)))

ggsave(paste0(MivsMo_save.dir, paste("MivsMo2_combined_severity_text_plots_all_lin.pdf"), sep=""),combined_plot, width = 6, height = 5)

## plot some variables in log scae
plot_list$B_19 <- plot_list$B_19 +scale_y_continuous(trans = 'log10') + annotation_logticks(sides="l")
plot_list$B_8 <- plot_list$B_8 +scale_y_continuous(trans = 'log10') + annotation_logticks(sides="l")

combined_plot <- do.call(ggarrange, c(plot_list, list(ncol = 3)))

ggsave(paste0(MivsMo_save.dir, paste("MivsMo2_combined_severity_text_plots_lin_log.pdf"), sep=""),combined_plot, width = 6, height = 5)

```


# Plots PRR
```{r}
vars_to_plot <- c("B_8" , "CD4_3", "Mono_4", "B_1", "B_9","B_19","NK_10","CD4_17")

sub_df <- flowsom_df_wide  %>%
  filter(comp3_PRR %in% "TRUE") %>%
  dplyr::select(subject_identifier,severity_text,timepoint,contains(c("CD4_","CD8_","B_","Mono_","NK_"))) %>%
  as.data.frame() %>% 
  mutate(severity_text = if_else(timepoint == 1, severity_text, "FALSE")) %>%
  mutate(severity_text = if_else(severity_text %in% c("mild", "moderate", "none"), "FALSE", as.factor(severity_text)))

sub_df$severe_case <- ifelse(sub_df$timepoint == 1 & sub_df$severity_text == "severe", "yes", "no")

# Loop through each variable and create a plot
for (var in vars_to_plot) {
  
 p <- ggplot(sub_df, aes(x = timepoint, y = !!sym(var), group = subject_identifier)) +
    geom_point(aes(color = severe_case, shape = severe_case, size = severe_case),
               show.legend = FALSE, alpha=0.5) +
    geom_line(aes(color = subject_identifier), size = 0.5, alpha=0.5) +
    geom_half_boxplot(aes(group = timepoint, fill = as.character(timepoint)), 
                      outlier.shape = NA, center = FALSE, errorbar.draw = FALSE, width = .4, 
                      position = position_nudge(-0.13)) +
    theme_classic() +
    theme(legend.position = "none") +
    scale_color_manual(values = c("no" = "black", "yes" = "red")) +
    scale_fill_manual(values = cols_TT_PRR) +
    scale_shape_manual(values = c("no" = 16, "yes" = 19)) +  # circle is 16, cross is 3
    scale_size_manual(values = c("no" = 1, "yes" = 2))  # change the size of points, 1 is for circle, 2 is for cross

  
  # Append the plot to the list
  plot_list[[var]] <- p
}

plot_list$B_18 <- plot_list$B_18 +scale_y_continuous(trans = 'log10') + annotation_logticks(sides="l")
plot_list$B_19 <- plot_list$B_19 +scale_y_continuous(trans = 'log10') + annotation_logticks(sides="l")

# Combine the plots using ggarrange
combined_plot <- do.call(ggarrange, c(plot_list, list(ncol = 3)))

# Save the combined plot as a single file
ggsave(paste0(PRR_save.dir, "combined_plots.pdf"), combined_plot, width = 10, height = 10)

```


# Networks
## adult vs children
```{r}
numeric_data_adult <- flowsom_df_wide %>% 
  filter(comp1_adultsvschildren_pos %in% "TRUE")%>%
  filter(adult_child %in% "adult") %>%
  dplyr::select(where(is.numeric)) %>% 
  dplyr::select(-c(age)) %>%
  select_if(~!all(. == 0)) %>%
  drop_na()

numeric_data_clean <- numeric_data_adult %>%
  select_if(~sd(.x) != 0)

cor_matrix_adult <- cor(numeric_data_clean)

group_labels <- sapply(strsplit(colnames(numeric_data_clean), "_"), "[[", 1)
group_indices <- lapply(unique(group_labels), function(x) which(group_labels == x))
names(group_indices) <- unique(group_labels)
group_indices_adult <- group_indices


numeric_data_child <- flowsom_df_wide %>% 
  filter(comp1_adultsvschildren_pos %in% "TRUE")%>%
  filter(adult_child %in% "child") %>%
  # filter(comp2_mildvsmoderate_adult %in% "TRUE")%>%
  dplyr::select(where(is.numeric)) %>% 
  dplyr::select(-c(age)) %>%
  select_if(~!all(. == 0)) %>%
  drop_na()

numeric_data_clean <- numeric_data_child %>%
  select_if(~sd(.x) != 0)

cor_matrix_child <- cor(numeric_data_clean)

group_labels <- sapply(strsplit(colnames(numeric_data_clean), "_"), "[[", 1)
group_indices <- lapply(unique(group_labels), function(x) which(group_labels == x))
names(group_indices) <- unique(group_labels)
group_indices_child <- group_indices
```

## mild, moderate, severe
```{r}
numeric_data_mild <- flowsom_df_wide %>% 
  filter(comp2_mildvsmoderate_adult %in% "TRUE")%>%
  filter(severity_text %in% "mild") %>%
  dplyr::select(where(is.numeric)) %>% 
  dplyr::select(-c(age)) %>%
  select_if(~!all(. == 0)) %>%
  drop_na()

numeric_data_clean <- numeric_data_mild %>%
  select_if(~sd(.x) != 0)

cor_matrix_mild <- cor(numeric_data_clean)

group_labels <- sapply(strsplit(colnames(numeric_data_clean), "_"), "[[", 1)
group_indices <- lapply(unique(group_labels), function(x) which(group_labels == x))
names(group_indices) <- unique(group_labels)
group_indices_mild <- group_indices


numeric_data_moderate <- flowsom_df_wide %>% 
  filter(comp2_mildvsmoderate_adult %in% "TRUE")%>%
  filter(severity_text %in% "moderate") %>%
  dplyr::select(where(is.numeric)) %>% 
  dplyr::select(-c(age)) %>%
  select_if(~!all(. == 0)) %>%
  drop_na()

numeric_data_clean <- numeric_data_moderate %>%
  select_if(~sd(.x) != 0)

cor_matrix_moderate <- cor(numeric_data_clean)

group_labels <- sapply(strsplit(colnames(numeric_data_clean), "_"), "[[", 1)
group_indices <- lapply(unique(group_labels), function(x) which(group_labels == x))
names(group_indices) <- unique(group_labels)
group_indices_moderate <- group_indices

numeric_data_severe <- flowsom_df_wide %>% 
  filter(comp2_mildvsmoderate_adult %in% "TRUE")%>%
  filter(severity_text %in% "severe") %>%
  dplyr::select(where(is.numeric)) %>% 
  dplyr::select(-c(age)) %>%
  select_if(~!all(. == 0)) %>%
  drop_na()

numeric_data_clean <- numeric_data_severe %>%
  select_if(~sd(.x) != 0)

cor_matrix_severe <- cor(numeric_data_clean)

group_labels <- sapply(strsplit(colnames(numeric_data_clean), "_"), "[[", 1)
group_indices <- lapply(unique(group_labels), function(x) which(group_labels == x))
names(group_indices) <- unique(group_labels)
group_indices_severe <- group_indices
```
## save adult child
```{r}
# Open a new PDF file
pdf("my_network_plots_AC_adult.pdf", height = 3.5, width = 3.5)
cor_network_adult <- qgraph(cor_matrix_adult, 
                            graph="cor",
                            layout = "spring", 
                            groups = group_indices_adult, 
                            palette="colorblind",
                            legend=FALSE, 
                            label.scale.equal=TRUE,
                            minimum="sig",
                            sampleSize=nrow(numeric_data_adult),
                            threshold=0.4,
                            label.scale = T,
                            vsize = 4)
dev.off()

pdf("my_network_plots_AC_child.pdf", height = 3.5, width = 3.5)
cor_network_child <- qgraph(cor_matrix_child, 
                            graph="cor",
                            layout = "spring", 
                            groups = group_indices_child, 
                            palette="colorblind",
                            legend=FALSE, 
                            label.scale.equal=TRUE,
                            minimum="sig",
                            sampleSize=nrow(numeric_data_child),
                            threshold=0.4,
                            label.scale = T,
                            vsize = 4)
dev.off()

pdf("my_network_plots_legend.pdf", height = 3.5, width = 3.5)
cor_network_child <- qgraph(cor_matrix_child, 
                            graph="cor",
                            layout = "spring", 
                            groups = group_indices_child, 
                            palette="colorblind",
                            legend=TRUE, 
                            label.scale.equal=TRUE,
                            minimum="sig",
                            sampleSize=nrow(numeric_data_child),
                            threshold=0.4,
                            label.scale = T,
                            vsize = 4)
dev.off()

```
## save mild moderate severe
```{r}
pdf("my_network_plots_mild.pdf", height = 3.5, width = 3.5)
cor_network_mild <- qgraph(cor_matrix_mild, 
                            graph="cor",
                            layout = "spring", 
                            groups = group_indices_mild, 
                            palette="colorblind",
                            legend=FALSE, 
                            label.scale.equal=TRUE,
                            label.cex=1,
                            minimum="sig",
                            sampleSize=nrow(numeric_data_mild),
                            threshold=0.4,
                            vsize = 4)
dev.off()

pdf("my_network_plots_moderate.pdf", height = 3.5, width = 3.5)
cor_network_moderate <- qgraph(cor_matrix_moderate, 
                            graph="cor",
                            layout = "spring", 
                            groups = group_indices_moderate, 
                            palette="colorblind",
                            legend=FALSE, 
                            label.scale.equal=TRUE,
                            label.cex=1,
                            minimum="sig",
                            sampleSize=nrow(numeric_data_moderate),
                            threshold=0.4,
                            vsize = 4)
dev.off()

pdf("my_network_plots_severe.pdf", height = 3.5, width = 3.5)
cor_network_severe <- qgraph(cor_matrix_severe, 
                            graph="cor",
                            layout = "spring", 
                            groups = group_indices_severe, 
                            palette="colorblind",
                            legend=FALSE, 
                            label.scale.equal=TRUE,
                            label.cex=1,
                            minimum="sig",
                            sampleSize=nrow(numeric_data_severe),
                            threshold=0.4,
                            vsize = 4)

# Close the PDF file
dev.off()
```

## Supplementary Figure 2
```{r}
pdf("network_adult_child_measure.pdf", height = 7, width = 6)
egg::ggarrange(

centralityPlot(cor_network_adult,include = 
               c("Strength","Betweenness","Closeness"),
               orderBy ="Strength"),
centralityPlot(cor_network_child,include = 
               c("Strength","Betweenness","Closeness"),
               orderBy ="Strength"),
ncol=2

)
dev.off()
```

## Supplementary Figure 3
```{r}
pdf("network_severity_measure.pdf", height = 10, width = 8)
egg::ggarrange(

centralityPlot(cor_network_mild,include = 
               c("Strength","Betweenness","Closeness"),
               orderBy ="Strength"),
centralityPlot(cor_network_moderate,include = 
               c("Strength","Betweenness","Closeness"),
               orderBy ="Strength"),
centralityPlot(cor_network_severe,include = 
               c("Strength","Betweenness","Closeness"),
               orderBy ="Strength"),
ncol=3

)
dev.off()
```

